---
title: Unified theming with brand.yml
description: Unified theming via brand.yml for Shiny and Quarto.
---

[brand.yml]: https://posit-dev.github.io/brand-yml
[quarto]: https://quarto.org
[shiny]: https://shiny.posit.co
[r markdown]: https://rmarkdown.rstudio.com



::: lead
Shiny for R now supports [brand.yml], a single YAML file that unifies visual theming across your projects. Define colors, fonts, and logos once in a YAML file, then use them consistently across Shiny apps, Quarto projects, and more.
:::

## What is brand.yml?

::: {.float-md-end .mx-3}
[![](brand-yml-tall.svg){alt="brand.yml logo" style="width: min(max(100%, 200px), 100%); max-height: 200px"}][brand.yml]{target="_blank"}
:::

[brand.yml] is a specification for managing visual branding through a single YAML file.
It simplifies brand management and theming by consolidating your visual identity---colors, typography, and styling---into one easy-to-maintain location.

At its most basic, a `_brand.yml` file includes metadata (`meta`) about the company or brand,
its logos (`logo`),
`color` palette, theme,
and the fonts and `typography` settings used by the brand.


Here's a simple `_brand.yml` file:

```{.yaml filename="_brand.yml"}
meta:
  name: brand.yml
  link: https://posit-dev.github.io/brand-yml

logo:
  small: brand-yml-icon.svg
  medium: brand-yml-tall.svg
  large: brand-yml-wide.svg

color:
  palette:
    orange: "#FF6F20"
    pink: "#FF3D7F"
  primary: orange
  danger: pink

typography:
  fonts:
    - family: Open Sans
      source: google
  base: Open Sans
```

This same file can be used to theme your Shiny apps as well as [Quarto] projects.
And brand.yml isn't just for corporate branding---you can also use brand.yml to create custom themes for your personal or team-specific projects.
Create a `_brand.yml` file to match a specific journal or conference style.

When used with bslib, simply place the `_brand.yml` file in your project directory, and bslib will automatically apply the branding to your Shiny app.

## Try brand.yml

bslib includes an example app that you can use to try out brand.yml features or that can be helpful as you develop your own `_brand.yml` file.
To run the app use `shiny::runExample()`.

```r
shiny::runExample("brand.yml", package = "bslib")
```

To use the app as a template, create a new directory for your own app and then copy the bslib example app to that directory

```{r eval=FALSE}
new_app_dir <- "my-brand" # path to your app
template <- fs::path_package("bslib", "examples-shiny/brand")

# Copy the template files into your new app directory
fs::dir_copy(template, new_app_dir)
```

Once created, open `app.R` and `_brand.yml` for editing.

```{r eval=FALSE}
file.edit(fs::path(new_app_dir, "app.R"))
file.edit(fs::path(new_app_dir, "_brand.yml"))
```

## Basic usage

### Automatic discovery

To use a [brand.yml] file like the example above, save the file as `_brand.yml` and include it in the directory containing your Shiny app.

```r
# Automatically find `_brand.yml` in app directory or its parent directories
ui <- page_fluid(
  theme = bs_theme()  # No additional configuration needed
)
```

When your app uses [uses `theme = bs_theme()`](../any-project/), any of the [`bslib::page_*()` functions](../dashboards/), or [bslib theming under-the-hood](../theming/), bslib will automatically

1. search for a `_brand.yml` file in the current directory or in a `_brand/` or `brand/` directory,
1. if not found, check parent directories for the `_brand.yml` file (possibly also in `_brand/` or `brand/`), and
2. apply the branded theming if a `_brand.yml` file is found.

### Explicit configuration

You can also be more explicit about brand.yml usage.

::: tabset

#### Ensure branding

Use `brand = TRUE` to find `_brand.yml` in app dir or its parents, or fail otherwise.

```r
ui <- page_sidebar(
  theme = bs_theme(brand = TRUE)
)
```

#### Direct path

Provide `brand` with a direct path to a brand.yml file, useful when the file has a special name or is stored in a central location.

```r
ui <- page_sidebar(
  theme = bs_theme(brand = "path/to/brand.yml")
)
```

#### Inline brand

Use a list to define brand settings inline, useful to quickly theme a small app.

```r
ui <- page_sidebar(
  theme = bs_theme(
    brand = list(
      color = list(primary = "#FF6F20", danger = "#FF3D7F"),
      typography = list(
        fonts = list(
          list(family = "Open Sans", source = "google")
        ),
        base = "Open Sans"
      )
    )
  )
)
```

### Ignore branding

Finally, you can ignore an existing `_brand.yml` file and apply no branding.

```r
ui <- page_sidebar(
  theme = bs_theme(brand = FALSE)
)
```

:::

### YAML configuration

Many places across the R Markdown ecosystem [use bslib for theming](../any-project/index.html#r-markdown).
Anywhere that `theme` is passed to bslib, you can use the same techniques mentioned above, even in YAML.

::: {.callout .callout-warning}
<h4 data-toc-skip>Bootstrap 5 Required</h4>

Theming with brand.yml works best with Bootstrap 5.
For best results, be sure to include `version: 5` under `theme`.
:::

::: tabset

#### Automatic

In the most basic case, place a `_brand.yml` in the project folder and ensure that the document uses Bootstrap 5.

```yaml
output:
  html_document:
    theme:
      version: 5
```

#### Ensure branding

Use `brand: TRUE` to automatically find `_brand.yml` or throw an error if not found within the project.

```yaml
output:
  html_document:
    theme:
      version: 5
      brand: true
```

#### Direct path

Provide `brand` with a direct path to a brand.yml file.

```yaml
output:
  html_document:
    theme:
      version: 5
      brand: "path/to/brand.yml"
```

#### Inline brand

Define brand settings directly inline in the `brand` key.

```yaml
output:
  html_document:
    theme:
      version: 5
      brand:
        color:
          primary: "#FF6F20"
          danger: "#FF3D7F"
        typography:
          fonts:
            - family: "Open Sans"
            source: "google"
          base: "Open Sans"
```

#### Ignore branding

Ignore an existing `_brand.yml` file and apply no branding with `brand: false`.

```yaml
output:
  html_document:
    theme:
      version: 5
      brand: false
```
:::

## Learn more

This article has barely scratched the surface of all of the information and customizations that brand.yml includes.
We encourage you to learn more about brand.yml and its features at the [official brand.yml website](https://posit-dev.github.io/brand-yml).
We also invite you to leave feedback and ask questions in the [brand.yml discussion board](https://github.com/posit-dev/brand-yml/discussions).

Visit these pages to learn more about using brand.yml with other projects:

1. [Quarto - Multiformat branding with _brand.yml](https://quarto.org/docs/authoring/brand.html)
1. [Shiny - Branded theming for Shiny for Python apps](https://shiny.posit.co/blog/posts/shiny-python-1.2-brand-yml/)

As a brand new project, most large language models (LLMs) don't know anything about brand.yml (yet), but they can still be useful with the right prompt.
We've created a small but complete description of brand.yml that you can use to help teach an LLM about brand.yml.
Read more in the [Creating a brand.yml with an LLM](https://posit-dev.github.io/brand-yml/articles/llm-brand-yml-prompt/) article on the [brand.yml] website.
