---
output: github_document
---

<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/bootstraplib)](https://cran.r-project.org/package=bootstraplib)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![R build status](https://github.com/rstudio/bootstraplib/workflows/R-CMD-check/badge.svg)](https://github.com/rstudio/bootstraplib/actions)
<!-- badges: end -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.align = "center"
)
library(bootstraplib)
```

# bootstraplib

The **bootstraplib** R package provides tools for creating custom [Bootstrap themes](https://getbootstrap.com/docs/4.4/getting-started/theming/), making it easier to style Shiny apps and R Markdown documents directly from R (via **sass**) without writing unruly CSS and HTML. Currently, **bootstraplib** supports Bootstrap 3 and 4, as well as a special `"4+3"` compatibility version, which allows you to start using Bootstrap 4 today in Shiny and R Markdown.

## Installation

**bootstraplib** isn’t yet available from CRAN, but you can install with:

``` r
remotes::install_github("rstudio/bootstraplib")
```

## Getting Started

### Create a theme

Use `bs_theme()` to create a **bootstraplib** theme, where you can:

* Choose a (major) Bootstrap version.
  * To current default is `version = "4+3"`, which means Bootstrap 4 plus an additional compatibility layer for Bootstrap 3 style navs, navbars, and more. This compatibility allows most Shiny apps and R Markdown documents to seamlessly upgrade to Bootstrap 4.
* Choose a Bootswatch theme (optional).
* Customize the main colors and fonts (see example below).
* More generally, customize any of Bootstrap's styling defaults via Sass variables.

For example, to implement a [material design inspired dark mode](https://material.io/design/color/dark-theme.html), set the main colors (and fonts):

```r
library(bootstraplib)
my_theme <- bs_theme(
  bg = "#202123", fg = "#B8BCC2", primary = "#EA80FC", 
  base_font = "Grandstander"
)
```

### Shiny usage

> Note: this usage requires the development version of Shiny `remotes::install_github("rstudio/shiny")`.

To use `my_theme` inside of Shiny, pass it to the relevant `theme` parameter in page functions such as `shiny::navbarPage()`, `shiny::fluidPage()`, `shiny::bootstrapPage()`, etc.

```r
library(shiny)
ui <- navbarPage(
  theme = my_theme,
  ...
)
shinyApp(ui, function(ui, server) {})
```

For a quick preview of how `my_theme` affects styling defaults for various "core" Shiny UI, use `bs_theme_preview()`. This preview app includes an interactive "real-time" theming widget for quickly iterating toward a set of desired set of colors and fonts. This theming widget emits code to replicate the styling changes in the R console and can also be added other Shiny apps via `run_with_themer()` (or `bs_themer()`).

```r
bs_theme_preview(my_theme)
```

```{r}
knitr::include_graphics("")
```

More generally, you can use a Bootstrap theme with any HTML page by using `shiny::bootstrapLib()` to provide the theme as an `htmltools::htmlDependency()` to an `shiny::htmlTemplate()` or any HTML `htmltools::tags` that you wish, for example:

```r
library(shiny)
ui <- htmlTemplate(
  "my-template.html",
  theme = my_theme,
  ...
)
shinyApp(ui, function(ui, server) {})
```

```html
<!-- my-template.html -->
<!DOCTYPE html>
<html>
  <head>
    {{ headContent() }}
    {{ bootstrapLib(theme) }}
  </head>
  <body>
    ...
  </body>
</html>
```

### R Markdown usage 

> Note: this usage currently requires an experimental version of R Markdown `remotes::install_github("rstudio/rmarkdown#1706")` and is subject to change

(1) Use `bootstrap_version` and `theme` to choose the Bootstrap version and a Bootswatch theme. These arguments are currently supported only in `html_document` and `html_document_base`.

```yaml
---
output:
  html_document:
    bootstrap_version: 4+3
    theme: minty
---
```

(2) Optionally add theme customizations inside any R code chunk (these customizations end up influencing the Bootstrap CSS included in the output document).

    ```{r}`r ''`
    library(bootstraplib)
    bs_global_theme_update(primary = 'green')
    ```

<!--
TODO: do we support passing of local theme objects?

```yaml
---
output:
  html_document:
    theme: `r bootstraplib::bs_theme(bg = "#202123", fg = "#B8BCC2", primary = "#EA80FC", "Grandstander")`
---
```
-->


## Learn more 

See the articles on [theming recipes](https://rstudio.github.io/bootstraplib/articles/recipes.html) and [foundations](https://rstudio.github.io/bootstraplib/articles/foundations.html).
