---
title: "Bootstrap 4 variables"
output: html_document
---

<script src="https://unpkg.com/@popperjs/core@2"></script>
<script src="https://unpkg.com/tippy.js@6"></script>

```{r setup, include = FALSE}
library(dplyr)
library(stringr)
library(purrr)
library(glue)
library(bslib)

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = FALSE, 
  message = FALSE
)

# Exclude all BS3 rules except for tables (for DT styling)
exclude <- c(
  '_normalize','_print', '_scaffolding', '_type', '_code',
  '_grid', '_forms', '_buttons', '_component-animations', '_dropdowns',
  '_button-groups', '_input-groups', '_navs', '_navbar', '_breadcrumbs',
  '_pagination', '_pager', '_labels', '_badges', '_jumbotron', '_thumbnails',
  '_alerts', '_progress-bars', '_media', '_list-group', '_panels', 
  '_responsive-embed', '_wells', '_close', '_modals', '_tooltip', '_popovers', 
  '_carousel', '_utilities', '_responsive-utilities'
)

bs_table_css <- bs_remove(bs_theme(version = 3), exclude)
bs_table_css <- bs_theme_dependencies(bs_table_css)[[2]]
bs_table_css$script <- NULL


htmltools::tagList(
  rmarkdown::html_dependency_font_awesome(),
  bs_table_css
)
```

```{r variables_df}
sass_variables_loc <- rprojroot::find_package_root_file("inst/lib/bs/scss/_variables.scss")
sass_variables_url <- file.path(
  "https://github.com/rstudio/bslib/blob", 
   substr(system("git rev-parse HEAD", intern = TRUE), 1, 7),
  "inst/lib/bs/scss/_variables.scss"
)
# A section is denoted by a comment followed by either an empty line or an empty comment.
# e.g.
# // Section ID
#
# Or
# // Section ID
# //
#
# Sometimes the sections are just isolated floating comments
#
# // Section ID
#
#
# TODO: BS3's comment templates are different so a whole new scraping pipeline is needed

raw_text <- brio::readLines(sass_variables_loc)

collapse_multiline_comments <- function(lines_df){

  repetitions <- rle(lines_df$is_comment & !lines_df$is_empty_comment)
  block_id <- rep(seq_along(repetitions$lengths), times = repetitions$lengths)
  is_comment_block <- rep(repetitions$values, times = repetitions$lengths)
  # A negative value means it's not a comment block. Using negatives so we
  # don't match ids of comment blocks
  lines_df$comment_block <- ifelse(is_comment_block, block_id, -seq_len(nrow(lines_df)))

  group_by(lines_df, comment_block) %>%
    summarise(
      across(c(-line), first),
      line = paste(line, collapse = " ")
    ) %>%
    arrange(line_number) %>%
    select(-comment_block)
}

link_to_pos_in_script <- function(line_num){
  paste0(sass_variables_url, "#L", line_num)
}

variables_df <- tibble(line = raw_text) %>%
  mutate(
    line_number = row_number(),
    is_comment = str_detect(line, "^\\/\\/"),
    is_empty_comment = str_detect(line, "^\\/\\/\\s*$")
  ) %>%
  collapse_multiline_comments() %>%
  mutate(
    below_is_empty_comment = lead(is_empty_comment),
    above_is_empty_line = lag(line) == "",
    below_is_empty_line = lead(line) == "",
    is_section = (is_comment & below_is_empty_comment) | (is_comment & above_is_empty_line & below_is_empty_line),
    variable = str_extract(line, "(?<=^\\$)(.+?)(?=:)"),
    section = ifelse(is_section, str_remove(line, "\\/\\/\\s*"), NA),
    section_link = ifelse(is_section, link_to_pos_in_script(line_number), NA),
    trailing_comment = str_extract(line, "(?<=;)(.+)$"),
    above_is_comment = lag(is_comment),
    comment = case_when(
      !is.na(trailing_comment) ~ trailing_comment,
      above_is_comment ~ lag(line),
      TRUE ~ ""
    ) %>% str_remove_all("\\/\\/") %>% str_trim(),
    comment = ifelse(str_detect(comment, "stylelint-disable"), "", comment)
  ) %>%
  tidyr::fill(section, section_link) %>%
  filter(!is.na(variable)) %>%
  mutate(
    value = str_extract(line, "(?<=:)(.+)?(?=;)") %>% str_remove(fixed("!default")) %>% str_trim(),
    value_is_function = str_detect(value, fixed("()")) | str_detect(value, "\\($"),
    # map-get function is used to reference a map built earlier so make sure not
    # to link to it because there is no entry in the table to link to.
    value = str_replace_all(
      value,
      "(?<!map-get\\()(\\$[\\w|-]+)",
      "<span class='dep-link'>\\1</span>"
    ),
    section = glue("Section: {section} <a href=\"{section_link}\"><i class=\"fas fa-external-link-alt\"></i></a>"),
    declaration = glue("<a href=\"{link_to_pos_in_script(line_number)}\"><i class=\"fas fa-external-link-alt\"></i></a>")
  ) %>%
  filter(!value_is_function) %>%
  select(
    section,
    variable,
    value,
    comment,
    declaration
  ) %>% 
  mutate(
    value = glue("<span data-tippy-content='CSS-value: {bs_get_variables(bs_theme(), variable)}'> {value} </span>")
  )
```

Below is a searchable table of Bootstrap 4 Sass variables. If you aren't sure what a Sass variable is or how to use them, see this [Get Started article](bslib.html#low-level). A few things to keep in mind when using this table:

* Each row represents a different Sass variable, showing its name and default value.
    * For Sass variables whose value defaults to another Sass variable (e.g., [`border-color`](#border-color)), you may hover over the value to get the (default) CSS value that it represents.
* Rows are grouped into sections of similar variables
    * For example, [Navs](#nav-link-padding-y) (`shiny::tabsetPanel()`), [Navbars](#navbar-padding-y) (`shiny::navbarPage()`), and [Dropdowns](#dropdown-min-width) (e.g., `shiny::navbarMenu()`).
* Towards the top of the table are more general theming options like `white`, `gray-*`, `black`, `primary`, `border-radius`, and so on, which end up impacting more specific theming variables like `btn-border-radius`.
    * `bs_theme()`'s `bg` and `fg` arguments provide a more convenient way to set the `white`, `gray-*`, and `black` variables, so there is no need to set these Sass variables directly (same goes for `base_font` -> `$font-family-base`, `heading_font` -> `$headings-font-family`, and `code_font` -> `$font-family-monospace`).


```{css}
table.dataTable tr[role="row"] td {
  font-family: monospace;
}
table.dataTable tr[role="row"] td:nth-child(3) {
  font-family: inherit;
  font-size: 0.85rem;
}
table.dataTable tr[role="row"] td:nth-child(4) {
  text-align: center;
}
table.dataTable tr td:nth-child(1) {
  max-width: 250px;
  overflow: scroll;
}
table.dataTable tr td:nth-child(2),
table.dataTable tr td:nth-child(3){
  max-width: 300px;
  overflow: scroll;
}
table.dataTable tr td:nth-child(4){
  width: 20px;
}
.dep-link {
  color:steelblue; 
  text-decoration:underline;
  cursor: pointer;
}
.selected .dep-link {
  color: white;
}
#table-wrapper {
  position: relative;
  --table_w: min(90vw, 1300px);
  width: var(--table_w);
  margin-left: calc(50% - var(--table_w)/2);
}
@media (min-width: 992px) {
  #table-wrapper.pkgdown {
    /* This is 65% instead of 50% because of how pkgdown/bootstrap 
       controls widths and the center content div is 75% of its 
       container div leaving an extra 15% to take into account */
    margin-left: calc(65% - var(--table_w)/2);
  }
}
```


```{js}
function find_and_select_variable(table, desired_var, update_page) {
  var row_index = table
    .column(1, { order: "current" })
    .data()
    .indexOf(desired_var);
  if (row_index !== -1) {
    table.rows({ selected: true }).deselect();
    table.row(row_index).select();
    location.hash = "#" + desired_var;

    if (update_page) {
      table.row(row_index).scrollTo();
    }
  }
}

function go_to_current_hash(table) {
  var desired_var = location.hash.replace("#", "");
  find_and_select_variable(table, desired_var, true);
}
```

<div id = "table-wrapper">
```{r}
variables_df %>%
  rename(
    Variable = variable,
    Value = value,
    Comment = comment,
    `Find in source` = declaration
  ) %>% 
  DT::datatable(
    escape = FALSE,
    style = "bootstrap",
    class = "cell-border",
    rownames = FALSE,
    selection = "none",
    extensions = c("Select", "RowGroup", "Scroller"),
    callback = DT::JS("
      // Check for linked variable on first load                    
      go_to_current_hash(table);
      
      // If this is a pkgdown environment we need to use different styling for table expansion
      if(document.querySelector('.template-article')){
        document.querySelector('#table-wrapper').classList.add('pkgdown');
      }
      // Allows us to link within the page
      window.onhashchange = function() {
        go_to_current_hash(table);
      }"),
    options = list(
      deferRender = TRUE,
      scroller = TRUE,
      scrollY = 850,
      scrollX = TRUE,
      rowGroup = list(dataSrc = 0),
      drawCallback = DT::JS("
        function( settings ) {
          var table = this.DataTable();
        
          tippy('[data-tippy-content]', {placement: 'right'});
        
          // Hide section column from view
          table.column(0).visible(false);
          
          // This makes variable names themselves selectable for building urls etc  
          this.find('tbody td:nth-child(1)').on('click', function(){
            find_and_select_variable(table, this.innerText, false);
          })
        
          // This makes the dependencies linked
          this.find('span.dep-link').on('click', function(){
            find_and_select_variable(table, this.innerText.replace('$', ''), true);
          });
        }")
    )
  )
```
</div>
