---
title: "Bootstrap 4 variables"
output: rmarkdown::html_vignette
author: "Nick Strayer & Carson Sievert"
date: "`r Sys.Date()`"
vignette: >
  %\VignetteIndexEntry{Bootstrap 4 variables}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<script src="https://unpkg.com/@popperjs/core@2"></script>
<script src="https://unpkg.com/tippy.js@6"></script>

```{r setup, include = FALSE}
library(dplyr)
library(stringr)
library(purrr)
library(glue)
library(bslib)

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = FALSE, 
  message = FALSE
)

htmltools::tagList(
  rmarkdown::html_dependency_font_awesome(),
  rmarkdown::html_dependency_bootstrap("default")
)
```

```{r variables_df}
sass_variables_loc <- rprojroot::find_package_root_file("inst/lib/bs/scss/_variables.scss")
sass_variables_url <- file.path(
  "https://github.com/rstudio/bslib/blob", 
   substr(system("git rev-parse HEAD", intern = TRUE), 1, 7),
  "inst/lib/bs/scss/_variables.scss"
)
# A section is denoted by a comment followed by either an empty line or an empty comment.
# e.g.
# // Section ID
#
# Or
# // Section ID
# //
#
# Sometimes the sections are just isolated floating comments
#
# // Section ID
#
#
# TODO: BS3's comment templates are different so a whole new scraping pipeline is needed

raw_text <- brio::readLines(sass_variables_loc)

collapse_multiline_comments <- function(lines_df){

  repetitions <- rle(lines_df$is_comment & !lines_df$is_empty_comment)
  block_id <- rep(seq_along(repetitions$lengths), times = repetitions$lengths)
  is_comment_block <- rep(repetitions$values, times = repetitions$lengths)
  # A negative value means it's not a comment block. Using negatives so we
  # don't match ids of comment blocks
  lines_df$comment_block <- ifelse(is_comment_block, block_id, -seq_len(nrow(lines_df)))

  group_by(lines_df, comment_block) %>%
    summarise(
      across(c(-line), first),
      line = paste(line, collapse = " ")
    ) %>%
    arrange(line_number) %>%
    select(-comment_block)
}

link_to_pos_in_script <- function(line_num){
  paste0(sass_variables_url, "#L", line_num)
}

variables_df <- tibble(line = raw_text) %>%
  mutate(
    line_number = row_number(),
    is_comment = str_detect(line, "^\\/\\/"),
    is_empty_comment = str_detect(line, "^\\/\\/\\s*$")
  ) %>%
  collapse_multiline_comments() %>%
  mutate(
    below_is_empty_comment = lead(is_empty_comment),
    above_is_empty_line = lag(line) == "",
    below_is_empty_line = lead(line) == "",
    is_section = (is_comment & below_is_empty_comment) | (is_comment & above_is_empty_line & below_is_empty_line),
    variable = str_extract(line, "(?<=^\\$)(.+?)(?=:)"),
    section = ifelse(is_section, str_remove(line, "\\/\\/\\s*"), NA),
    section_link = ifelse(is_section, link_to_pos_in_script(line_number), NA),
    trailing_comment = str_extract(line, "(?<=;)(.+)$"),
    above_is_comment = lag(is_comment),
    comment = case_when(
      !is.na(trailing_comment) ~ trailing_comment,
      above_is_comment ~ lag(line),
      TRUE ~ ""
    ) %>% str_remove_all("\\/\\/") %>% str_trim(),
    comment = ifelse(str_detect(comment, "stylelint-disable"), "", comment)
  ) %>%
  tidyr::fill(section, section_link) %>%
  filter(!is.na(variable)) %>%
  mutate(
    value = str_extract(line, "(?<=:)(.+)?(?=;)") %>% str_remove(fixed("!default")) %>% str_trim(),
    value_is_function = str_detect(value, fixed("()")) | str_detect(value, "\\($"),
    # map-get function is used to reference a map built earlier so make sure not
    # to link to it because there is no entry in the table to link to.
    value = str_replace_all(
      value,
      "(?<!map-get\\()(\\$[\\w|-]+)",
      "<span class='dep-link'>\\1</span>"
    ),
    section = glue("Section: {section} <a href=\"{section_link}\"><i class=\"fas fa-external-link-alt\"></i></a>"),
    declaration = glue("<a href=\"{link_to_pos_in_script(line_number)}\"><i class=\"fas fa-external-link-alt\"></i></a>")
  ) %>%
  filter(!value_is_function) %>%
  select(
    section,
    variable,
    value,
    comment,
    declaration
  ) %>% 
  mutate(
    value = glue("<span data-tippy-content='CSS-value: {bs_get_variables(bs_theme(), variable)}'> {value} </span>")
  )
```

Not only does `bs_theme()` make it easy to control [Main Colors & Fonts](theming.html#main-colors), but it also provides access to 100s of more specific theming options. These options, sometimes referred to as [Bootstrap Sass variables](https://getbootstrap.com/docs/4.4/getting-started/theming/#variable-defaults), allow for control over default styles of various Bootstrap features like [Navs](#nav-link-padding-y) (e.g., `shiny::tabsetPanel()`), [Navbars](#navbar-padding-y) (e.g., `shiny::navbarPage()`), [Dropdowns](#dropdown-min-width) (e.g., `shiny::navbarMenu()`), and much more.

The table below provides a convenient interface for discovering and learning about Bootstrap 4 Sass variables. Here are a few things to keep in mind when using this table:

1. Towards the top of the table are more general theming options like `white`, `gray-*`, `black`, `primary`, `border-radius`, and so on, which end up impacting more specific theming variables like `btn-border-radius`.
2. `bs_theme()`'s `bg` and `fg` arguments provide a more convenient way to set the `white`, `gray-*`, and `black` variables, so there is no need to set these Sass variables directly (same goes for `font_base` -> `$font-family-base`, `font_heading` -> `$font-family-base`, and `font_code` -> `$font-family-monospace`).
3. For Sass variables whose value defaults to another Sass variable (e.g., [`border-color`](bs4-variables.html#border-color)), you may hover over the value to get the (default) CSS value that it represents.

For more information and examples on why and how to use these Sass variables inside of `bs_theme()`, see the [theming variables section](theming.html#sass-variables) of the Custom Theming article.

```{css}
table.dataTable tr[role="row"] td {
  font-family: monospace;
}
table.dataTable tr[role="row"] td:nth-child(3) {
  font-family: inherit;
  font-size: 0.85rem;
}
table.dataTable tr[role="row"] td:nth-child(4) {
  text-align: center;
}
table.dataTable tr td:nth-child(1) {
  max-width: 250px;
  overflow: scroll;
}
table.dataTable tr td:nth-child(2),
table.dataTable tr td:nth-child(3){
  max-width: 300px;
  overflow: scroll;
}
table.dataTable tr td:nth-child(4){
  width: 20px;
}
.dep-link {
  color:steelblue; 
  text-decoration:underline;
  cursor: pointer;
}
.selected .dep-link {
  color: white;
}

#table-wrapper {
  position: relative;
  width: 90vw;
  margin-left: calc(50% - 45vw);
}
  

@media (min-width: 992px) {
  #table-wrapper.pagedown {
    /* This is 65% instead of 50% because of how pagedown/bootstrap 
       controls widths and the center content div is 75% of its 
       container div leaving an extra 15% to take into account */
    margin-left: calc(65% - 45vw);
  }
}
```


```{js}
function find_and_select_variable(table, desired_var, update_page) {
  var row_index = table
    .column(1, { order: "current" })
    .data()
    .indexOf(desired_var);
  if (row_index !== -1) {
    table.rows({ selected: true }).deselect();
    table.row(row_index).select();
    location.hash = "#" + desired_var;

    if (update_page) {
      table.row(row_index).scrollTo();
    }
  }
}

function go_to_current_hash(table) {
  var desired_var = location.hash.replace("#", "");
  find_and_select_variable(table, desired_var, true);
}


```

<div id = "table-wrapper">
```{r}
variables_df %>%
  rename(
    Variable = variable,
    Value = value,
    Comment = comment,
    `Find in source` = declaration
  ) %>% 
  DT::datatable(
    escape = FALSE,
    style = "bootstrap",
    class = "cell-border",
    rownames = FALSE,
    selection = "none",
    extensions = c("Select", "RowGroup", "Scroller"),
    callback = DT::JS("
      // Check for linked variable on first load                    
      go_to_current_hash(table);
      
      // If this is a pagedown environment we need to use different styling for table expansion
      if(document.querySelector('.template-article')){
        document.querySelector('#table-wrapper').classList.add('pagedown');
      }
      // Allows us to link within the page
      window.onhashchange = function() {
        go_to_current_hash(table);
      }"),
    options = list(
      deferRender = TRUE,
      scroller = TRUE,
      scrollY = 850,
      scrollX = TRUE,
      rowGroup = list(dataSrc = 0),
      drawCallback = DT::JS("
        function( settings ) {
          var table = this.DataTable();
        
          tippy('[data-tippy-content]', {placement: 'right'});
        
          // Hide section column from view
          table.column(0).visible(false);
          
          // This makes variable names themselves selectable for building urls etc  
          this.find('tbody td:nth-child(1)').on('click', function(){
            find_and_select_variable(table, this.innerText, false);
          })
        
          // This makes the dependencies linked
          this.find('span.dep-link').on('click', function(){
            find_and_select_variable(table, this.innerText.replace('$', ''), true);
          });
        }")
    )
  )
```
</div>
