---
title: "Filling layouts"
resource_files:
  - examples/filling
---

```{r setup, include = FALSE}
library(plotly)
source(
  rprojroot::find_package_root_file("vignettes/common.R")
)
knitr::opts_chunk$set(echo = FALSE)
```

```{scss, echo = FALSE}
.section.level2 {
  margin-top: 2rem;
}
.section.level3 {
  margin-top: 1rem;
}
```

This article covers the building blocks of filling layouts in `bslib`: **fillable containers** and **fill items**. Filling layout is an inherently nuanced topic, and fully groking takes some effort, but you'll gain a powerful way to get UI elements to fill the window, fill inside [cards](cards.html), fill inside [sidebar layouts](sidebars.html), or generally fill anywhere you want.

Since, in theory, essentially any UI element can be coerced into a **fillable container** and/or **fill item**[^flexbox], it's useful to first study their behavior in the abstract, which we do next in the [In theory](#in-theory) section. After, the [In practice](#in-practice) section reinforces those concepts and demonstrates their power with practical examples.

[^flexbox]: Technically speaking, a **fillable container** is just a [CSS flexbox](https://css-tricks.com/snippets/css/a-guide-to-flexbox/) container with `flex-direction: column` (which makes it's children flex items). And, when a **fill item** appears as a direct child of a **fillable container**, it is given a `flex` property of `1` (i.e., it's allowed to grow/shrink).

Throughout these sections, be aware that most `bslib` components, as well as many Shiny outputs (e.g., `plotOutput()`, `plotlyOutput()`, etc) classify as [fill items]{.b-blue} by default. This means they possess the potential to grow/shrink to fit their container, but that potential is only activated when their _immediate parent_ is a [fillable container]{.b-pink} _with a defined height_. Also be aware that `bslib` components and many Shiny outputs have `fill` and/or `fillable` arguments to opt out/in of this behavior (and `bslib` also provides an API for testing/coercing these properties on _any_ UI element -- see `is_fill()` for more).

## In theory

```{r setup-functions, echo=FALSE}
fillable_container <- function(..., tooltip = NULL, resizable = FALSE, padding = TRUE) {
  gap <- "0.25rem"
  card(
    "data-bs-toggle" = if (!is.null(tooltip)) "tooltip",
    "data-bs-title" = paste(tooltip, collapse = "<br>"),
    "data-bs-html" = "true",
    "data-bs-placement" = "top",
    "data-bs-custom-class" = "tip-pink",
    style = css(
      "--bs-card-border-color" = "var(--bs-pink)",
      "--bs-card-border-width" = "0.12rem",
      gap = if (padding) gap,
      resize = if (resizable) "vertical"
    ),
    ...,
    wrapper = function(...) {
      card_body(
        class = if (padding) "p-1" else "p-0",
        gap = if (padding) gap,
        ...
      )
    }
  )
}


child <- function(..., fill = TRUE, fillable = FALSE, tooltip = NULL, placement = "right", height = 400) {
  both <- fill && fillable
  color <- if (both) "purple" else if (fill) "blue" else "orange"
  div(
    class = "border border-1 border-dark rounded text-center",
    class = paste0("bg-fill-", color),
    "data-bs-toggle" =  if (!is.null(tooltip)) "tooltip",
    "data-bs-title" = paste(tooltip, collapse = "<br>"),
    "data-bs-html" = "true",
    "data-bs-placement" = placement,
    "data-bs-custom-class" = paste0("tip-", color),
    style = css(
      color = "white",
      "--bs-tooltip-bg" = sprintf("var(--bs-%s)", color),
      height = validateCssUnit(height),
      flex_shrink = if (!fill) 0,
      padding = "0.2rem"
    ),
    ...
  ) |>
    bindFillRole(item = fill, container = fillable)
}
```

```{scss, echo=FALSE}
.tip-pink {
  --bs-tooltip-bg: var(--bs-pink);
}
.tip-blue {
  --bs-tooltip-bg: var(--bs-blue);
}
.tip-orange {
  --bs-tooltip-bg: var(--bs-orange);
}
.tip-purple {
  --bs-tooltip-bg: var(--bs-purple);
}

.b-pink {
	font-weight: bold;
	color: var(--bs-pink);
}
.b-orange {
	font-weight: bold;
	color: var(--bs-orange);
}
.b-blue {
	font-weight: bold;
	color: hsl(202deg, 100%, 38%);
}
.b-purple {
  font-weight: bold;
  color: var(--bs-purple);
}

.bg-fill-blue {
	background-image: linear-gradient(
    180deg,
    hsl(202deg, 100%, 38%) 0%,
    hsl(206deg, 58%, 55%) 25%,
    hsl(206deg, 58%, 66%) 50%,
    hsl(205deg, 61%, 77%) 75%,
    hsl(203deg, 69%, 87%) 100%
  );
}

.bg-fill-orange {
  background-image: linear-gradient(
    180deg,
    hsl(27deg, 98%, 54%) 0%,
    hsl(27deg, 97%, 64%) 25%,
    hsl(27deg, 93%, 73%) 50%,
    hsl(28deg, 79%, 82%) 75%,
    hsl(28deg, 27%, 90%) 100%
  );
}

.bg-fill-purple {
  background: var(--bs-purple);
}

```


### Activating fill

::: row
::: col-md-6
Just like any other HTML **container**, a [fillable container]{.b-pink}'s default height depends on the height of it's children. So, for example, if there's a single [fill item]{.b-blue} with a defined height of `400px` (the default for most Shiny outputs), the [fillable container]{.b-pink}'s height is also `400px` (plus any padding, border, etc).
:::

::: col-md-6
```{r}
fillable_container(
  tooltip = c("Fillable container", "Computed height: 400px"),
  child(
    tooltip = c("Fill child", "Defined height: 400px")
  )
)
```
:::
:::


::: {.row .mt-3}
::: col-md-6
Defining the height of a [fillable container]{.b-pink} activates its immediate children's potential to [fill]{.b-blue}. So, for example, if [fillable container]{.b-pink}'s height is set to `200px`, the [fill child]{.b-blue} would shrink to about `200px`:
:::

::: col-md-6
```{r}
fillable_container(
  tooltip = c("Fillable", "Defined height: 200px"),
  child(
    tooltip = c("Fill", "Defined height: 400px", "Computed height: 200px")
  ),
  height = 200
)
```
:::
:::


::: {.row .mt-3}
::: col-md-6
If multiple [fill items]{.b-blue} were immediate children of this [fillable container]{.b-pink}, they'd keep shrinking (in this case, to about `100px` each):
:::

::: col-md-6
```{r}
fillable_container(
  tooltip = c("Fillable", "Defined height: 200px"),
  child(
    tooltip = c("Fill", "Defined height: 400px", "Computed height: 100px")
  ), 
  child(
    tooltip = c("Fill", "Defined height: 400px", "Computed height: 100px")
  ),
  height = 200
)
```
:::
:::


::: {.row .mt-3}
::: col-md-6
Adding a [non-fill]{.b-orange} item (e.g., `htmltools::p()`-aragraph of text) won't cause that particular item to grow/shrink, but the [fill]{.b-blue} items divvy up any remaining space (**careful:** if [non-fill]{.b-orange} item(s) are larger than the [fillable container]{.b-pink}, the [fill]{.b-blue} items won't be visible!). This is big reason why `card()`s have a `min_height` argument (to prevent [fill]{.b-blue} items from shrinking too much).
:::

::: col-md-6
```{r}
fillable_container(
  tooltip = c("Fillable container"),
  child(tooltip = "Fill"), 
  child(tooltip = "Fill"),
  child(
    fill = FALSE, height = 100,
    tooltip = "Non-fill"
  ), 
  height = 200,
  resizable = TRUE
)
```
:::
:::


::: {.callout .callout-warning}
<h3 data-toc-skip>Resizable example</h3>


Notice the resizing handle on the lower-right hand corner of the [fillable container]{.b-pink} above. Use it to change the size of the [fillable container]{.b-pink} and compare the behavior between [fill]{.b-blue} and [non-fill]{.b-orange} items.
:::


### Carrying fill

The previous section focuses on the fairly simple case of _one_ parent container. However, in practice, you'll likely be working with multiple levels of parents, which quickly complicates things, especially because:

1. [Fill items]{.b-blue} require their _immediate_ parent to be a [fillable container]{.b-pink} in order to fill.
2. All "raw" HTML tags (e.g., `div()`, `p()`, etc.) as well as many Shiny UI elements (e.g., `wellPanel()`, etc.) are neither [fillable]{.b-pink} nor [fill]{.b-blue} (i.e., we'll call these [non-fill]{.b-orange} elements).


::: {.row .mt-3}
::: col-md-6
As a result, a common way in which (1) breaks down is that a [non-fill]{.b-orange} element, like a `div()`, comes between [fillable]{.b-pink} and [fill]{.b-blue}. In fact, you'll run into this exact behavior when using `uiOutput()` to insert a dynamically rendered [fill]{.b-blue} item into a [fillable container]{.b-pink} (see [this section](#ui-output) for a concrete example).
:::

::: col-md-6
```{r}
fillable_container(
  tooltip = "Fillable container",
  child(
    fill = FALSE, 
    tooltip = "Non-fill",
    "data-bs-offset" = "-140,0",
    child(
      tooltip = "Fill",
      "data-bs-offset" = "-100,0"
    )
  ),
  height = 200,
  resizable = TRUE
)
```
:::
:::

::: {.row .mt-3}
::: col-md-6
Assuming the goal is for the [fill item]{.b-blue} to fit the [fillable container]{.b-pink}, it's useful to coerce the [non-fill]{.b-orange} element into both [fill item]{.b-blue} _and_ a [fillable container]{.b-pink}, which we call a [fill carriers]{.b-purple}. Any UI element can be coerced into a [fill carrier]{.b-purple} with `as_fill_carrier()`.
:::

::: col-md-6
```{r}
fillable_container(
  tooltip = "Fillable container",
  child(
    fillable = TRUE,
    tooltip = "Fill carrier",
    "data-bs-offset" = "-40,0",
    child(tooltip = "Fill")
  ),
  height = 200,
  resizable = TRUE
)
```
:::
:::


::: {.row .mt-3}
::: col-md-6
This concept of a [fill carrier]{.b-purple} is especially useful and relevant for [cards](cards.html). In most cases, a [card]{.b-pink} has numerous children like a [header]{.b-orange} and a [body]{.b-purple}, and the body commonly contains [fill item]{.b-blue}(s) (to ensure [fill item]{.b-blue}s). This is why `card_body()`{.b-purple} defaults to `fillable = TRUE` (and `fill = TRUE`).
:::

::: col-md-6
```{r}
fillable_container(
  tooltip = c("<code>card()</code> (fillable container)"),
  child(
    fill = FALSE, height = 50,
    tooltip = "<code>card_header()</code>"
  ), 
  child(
    fillable = TRUE,
    tooltip = "<code>card_body()</code>",
    child(
      tooltip = "Fill",
      "data-bs-offset" = "-30,0"
    )
  ),
  height = 200,
  resizable = TRUE
)
```
:::
:::

::: {.row .mt-3}
::: col-md-6
You might wonder, why then would we want or need `fillable = FALSE` or `fill = FALSE` on a `card_body()`? One big reason is that [fillable containers]{.b-pink} are powered by [CSS flexbox](https://css-tricks.com/snippets/css/a-guide-to-flexbox/), which changes the way it's children are rendered. And, although those changes are nice for "stretchy" children, there are [downsides for rendering inline elements](cards.html#flexbox). So, that's why, it's recommended that you use [multiple card bodies](#multiple-card_body) when combining [fill]{.b-blue} with [non-fill]{.b-orange}
:::

::: col-md-6
```{r}
fillable_container(
  tooltip = c("<code>card()</code> (fillable container)"),
  child(
    fill = FALSE, height = 50,
    tooltip = "<code>card_header()</code> (not fill)"
  ), 
  child(
    fillable = TRUE, 
    tooltip = "<code>card_body()</code>",
    child(
      tooltip = "Fill",
      "data-bs-offset" = "-30,0"
    )
  ),
  child(
    fill = FALSE, height = 40,
    tooltip = "<code>card_body(fillable = F, fill = F)</code>",
    placement = "top",
    child(
      fill = FALSE, height = 30, 
      tooltip = "Not fill",
    )
  ),
  height = 300,
  resizable = TRUE
)
```
:::
:::


## In practice

This section puts into practice what we learned in [the theory](#in-theory) of [fillable containers]{.b-pink} and [fill items]{.b-blue}.

```{r, include = FALSE}
opts_chunk$set(echo = TRUE)
```

### Setup code

The example in the sub-sections that follow assume you've ran the following code. Here we're using `{plotly}` to create a list of [fill items]{.b-blue}, but the same concepts extend to other htmlwidgets (e.g., `{leaflet}`) and Shiny outputs like `plotOutput()`.[^troubleshoot]

[^troubleshoot]: If these techniques don't extend as advertised to your output(s) of choice, see the [other advice](#other-advice) section to help troubleshoot. If that doesn't help, please [let us know about it](https://github.com/rstudio/bslib/issues/new/choose)!

```{r}
library(plotly)
plots <- list(
  plot_ly(diamonds) |> add_histogram(x = ~price),
  plot_ly(diamonds) |> add_histogram(x = ~carat),
  plot_ly(diamonds) |> add_histogram(x = ~cut, color = ~clarity)
)
plots <- lapply(plots, function(x) {
  config(x, displayModeBar = FALSE) |>
    layout(margin = list(t = 0, b = 0, l = 0, r = 0))
})
```

### Filling the window

Perhaps the most important [fillable container]{.b-pink} is `page_fillable()`, which sets it's height equal to the browser window. Thus, if [fill items]{.b-blue} appear as direct children, they'll fill the window. `page_fillable()` also defaults to `fillable_mobile = FALSE`, which means the height isn't set equal to the viewport on mobile. As a result, fill items use their own defined height (instead of the viewport size) on mobile, which is often better behavior when showing multiple outputs.


```{r as_iframe=TRUE, resizable=TRUE}
page_fillable(
  h2("Diamond plots"),
  plots[[1]], plots[[2]], plots[[3]]
)
```

:::: {.row .mt-3}

::: {.col-md-6 .mt-auto .mb-auto}
::: {.callout .callout-note}
<h3 data-toc-skip>Resizable example</h3>

Notice the resizing handle on the lower-right hand corner of the example above. Use it to change the size of the "window" and see the behavior of the filling plots  
:::
:::

::: col-md-6
::: {.callout .callout-note}
<h3 data-toc-skip>Limiting shrinkage</h3>

If you're worried about plots becoming too small, consider putting them in a `card_body()` with a `min_height` (like we do later on). Also, if you don't want the card border, you can do `card(class = "border-0", ...)`
:::
:::
::::

### Multiple columns

Since `layout_columns()` is a [fill item]{.b-blue} (by default), it grows/shrinks just like any other [fill item]{.b-blue}. It also defaults to `fillable = TRUE`, which in this case, means each column gets wrapped in a [fillable container]{.b-pink}. That's why, in this example, `plots[[1]]` and `plots[[1]]` also grow/shrink to match the size of the `layout_columns()` container.

```{r as_iframe=TRUE, resizable=TRUE}
page_fillable(
  h2("Diamond plots"),
  layout_columns(plots[[1]], plots[[2]]),
  plots[[3]]
)
```

### Value boxes

Since `value_box()` is a [fill item]{.b-blue} (by default), it grows/shrinks just like any other [fill item]{.b-blue}. This is especially useful for keeping a common baseline in a multi-column layout. That said, the multi-layout column that holds value boxes probably doesn't want it default `fill = TRUE` behavior, since the value boxes should be given more/less space and the window becomes larger/smaller:

```{r as_iframe=TRUE, resizable=TRUE}
boxes <- layout_columns(
  fill = FALSE,
  value_box(
    "Total diamonds",
    scales::comma(nrow(diamonds)),
    showcase = bsicons::bs_icon("gem", size = NULL)
  ),
  value_box(
    "Average price", 
    scales::dollar(mean(diamonds$price), accuracy = 1),
    showcase = bsicons::bs_icon("coin", size = NULL),
    theme_color = "success"
  ),
  value_box(
    "Average carat",
    scales::number(mean(diamonds$carat), accuracy = .1),
    showcase = bsicons::bs_icon("search", size = NULL),
    theme_color = "dark"
  )
)

page_fillable(
  boxes,
  layout_columns(plots[[1]], plots[[2]]),
  plots[[3]]
)
```

::: {.callout .callout-note}
<h3 data-toc-skip>Column wrapping layouts</h3>

To learn more about `layout_columns()`, see [this article](column-layout.html).
:::


### Full-screen cards

As alluded to in the [Carrying fill](#carrying-fill) section, `card()` and `card_body()` are [fill carriers]{.b-purple} (that is, they are both [fillable]{.b-pink} and [fill]{.b-blue}, by default). Therefore, by wrapping each plot in a card, the card not only grows/shrinks (since they are [fill]{.b-blue}), but also retain the plot's ability to grow/shrink (since they are [fillable]{.b-pink}).

```{r as_iframe=TRUE, resizable=TRUE}
plot_card <- function(header, ...) {
  card(
    full_screen = TRUE,
    card_header(header, class = "bg-dark"),
    card_body(..., min_height = 150)
  )
}

page_fillable(
  layout_columns(
    plot_card("Diamond price", plots[[1]]), 
    plot_card("Diamond carat", plots[[2]])
  ),
  plot_card("Diamond cut by clarity", plots[[3]])
)
```

Note that, if we changed `page_fillable()` to `page_fluid()` (or `page_fixed()`), each plot would render to it's default height (`400px`) since we no longer have a [fillable]{.b-pink} with a specified height. That said, even in that case, if we expand the card to full-screen, the plot still grows to fit the full screen card (since the `card()` is then a [fillable]{.b-pink} container with a specified height, the `card_body()` is a [fill carrier]{.b-purple}, and the plot is a [fill item]{.b-blue}).


### Sidebar layouts

Similar to what we've seen with outputs and `card()`s, `layout_sidebar()` is also a fill item (by default), so placing it as a direct child of `page_fillable()` makes it fit the window. Also, the main content's container defaults to a fillable container, so if that behavior is undesirable, set `fillable = FALSE` in `layout_sidebar()`.

```{r as_iframe=TRUE, resizable=TRUE}
page_fillable(
  padding = 0,
  layout_sidebar(
    border = FALSE,
    fillable = FALSE,
    sidebar = sidebar(
      title = "Diamond plots",
      "Input controls here..."
    ),
    layout_columns(
      plot_card("Diamond price", plots[[1]]), 
      plot_card("Diamond carat", plots[[2]])
    ),
    plot_card("Diamond cut by clarity", plots[[3]])
  )
)
```

::: {.callout .callout-note}
<h3 data-toc-skip>Sidebar layouts</h3>

To learn more about `layout_sidebar()`, see [this article](sidebars.html).
:::


## Other advice

### Dynamic UI {#ui-output}

:::: {.row .mt-3}

::: col-md-6
As alluded to in the [Carrying fill](#carrying-fill) section, `uiOutput()` puts an additional UI element around `renderUI()`'s return value. So, in order to carry the potential to fill down to a fill item (e.g., `plot_ly()`), mark `uiOutput()` as a fill carrier.
:::

::: col-md-6
```{r, eval = FALSE}
library(plotly)

ui <- page_fluid(
  card(
    full_screen = TRUE,
    max_height = 300,
    card_header("My plot"),
    uiOutput("plot", as_fill_carrier())
  )
)

server <- function(input, output) {
  output$plot <- renderUI({
    plot_ly(diamonds, x = ~price)
  })
}

shinyApp(ui, server)
```
:::
::::

### DT tables

:::: {.row .mt-3}

::: col-md-6
`{DT}`'s `datatable()` has it's own unique interface for filling a container. Specifically, make sure to set `datatable(fillContainer = TRUE)` in order for the table to grow/shrink as you'd expect it to.
:::

::: col-md-6
```{r, eval = FALSE}
library(DT)

ui <- page_fluid(
  card(
    full_screen = TRUE,
    max_height = 350,
    card_header("My table"),
    dataTableOutput("dt")
  )
)

server <- function(input, output) {
  output$dt <- renderDataTable({
    datatable(
      mtcars, fillContainer = TRUE
    )
  })
}

shinyApp(ui, server)
```
:::
::::

### Other htmlwidgets

Broadly speaking, most `htmlwidgets` like `{plotly}` and `{leaflet}` are [fill items]{.b-blue} by default, but that might not always be the case. Also, sometimes, you might not want a particular widget to be treated as a [fill item]{.b-blue}. In the Shiny case, you should be able to control this through a `fill` argument on the output container (e.g., `plotlyOutput("id", fill = FALSE)`), but if no `fill` argument is available you can also use `bslib`'s `as_fill()` API to opt in/out. In the non-Shiny case, you can control `fill` through the widget's `htmlwidgets::sizingPolicy()` (e.g., `leaflet()$sizingPolicy$fill`).


### Avoid `fluidRow()`/`column()`

Modern versions of [Bootstrap Grid](https://getbootstrap.com/docs/5.3/layout/grid/) currently use [CSS Flexbox](https://css-tricks.com/snippets/css/a-guide-to-flexbox/) in such a way that filling layout is mostly incompatible with `fluidRow()`/`column()`. Instead, use `layout_columns()` to implement [multi-column filling layouts](column-layout.html)


<!-- Extra js bits --->

```{=html}
<script>
const tooltipTriggerList = document.querySelectorAll('[data-bs-toggle=\"tooltip\"]')
const tooltipList = [...tooltipTriggerList].map(el => {
  new bootstrap.Tooltip(el, {
    trigger: 'hover'
  })
})
</script>
```
