---
title: "Rmarkdown & bootstraplib"
author: "Carson Sievert"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Rmarkdown \& bootstraplib}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Bootstrap 4

> **NOTE**: This functionality currently requires installing an experimental branch of **rmarkdown**: `remotes::install_github('rstudio/rmarkdown#1706')`.

To use Bootstrap 4 in **rmarkdown**, use the `bootstrap_version` parameter to the `html_document` output format. In addition to this "strict" Bootstrap 4 mode, there is "compatible" Bootstrap 4 mode (i.e., `bootstrap_version: 4-3`), which allows you to rely on BS3 style nav/navbar markup. Only use this compatibility mode if you have existing `html_document`/`html_document_base` output formats that rely on BS3 markup.

```yaml
---
output:
  html_document:
    bootstrap_version: 4
---
```

### Bootswatch themes

Regardless of the `bootstrap_version`, a Bootswatch theme may be added via the `theme` parameter. When using Bootstrap 4, you can use some new Bootswatch themes (e.g., sketchy, minty, lux, and solar).

```yaml
---
output:
  html_document:
    bootstrap_version: 4
    theme: sketchy
---
```

Here's a list of all Bootswatch 4 themes:

```{r}
bootstraplib::bootswatch_themes(4)
```

### Custom themes

In **rmarkdown** documents, you don't need to [create a new theme](#custom-create) to start [adding to it](#custom-add), because the `bootstrap_version` and `theme` parameters are used to create one for you. Thus, to start customizing the styling of a given document, use `bs_theme_add_variables()` and/or `bs_theme_add()` inside any R code chunk ([see below](#custom-add) for more).

    ```{r}`r ''`
    library(bootstraplib)
    bs_theme_add_variables(primary = 'salmon')
    ```

See the [custom themes vignette](custom.html) to learn more.
