```{r setup, include = FALSE}
## Uncomment the line below and knit this document to create or update the
## screenshots of the navigation containers.
# SCREENSHOT <- TRUE
```

```{r echo = FALSE, eval = !exists("SCREENSHOT"), include = FALSE}
render_screenshot_img <- function(x, options) {
  file_name <- paste0(options$label, ".png")
  alt <- options$fig.alt
  if (is.null(alt)) {
    alt <- sprintf("Screenshot of a %s() example.", options$label)
  }
  img <- sprintf('![](%s "%s")', file_name, alt)
  knitr::asis_output(img)
}
knitr::opts_chunk$set(eval = TRUE, render = render_screenshot_img)
```

```{r eval = exists("SCREENSHOT"), include = FALSE}
# take screenshots of each nav panel setup
library(chromote)
library(bslib)
b <- ChromoteSession$new(width = 800, height = 300)

render_screenshot <- function(x, options) {
  if (inherits(x, "bslib_fragment")) {
    x <- attr(x, "bslib_page")(x)
  }
  x <- tagList(
    x,
    tags$head(tags$style("
      body { min-height: 300px; padding: 1rem }
      .bslib-card .card-body { min-height: 200px; }
    "))
  )
  tmpdir <- tempfile()
  dir.create(tmpdir)
  on.exit(unlink(tmpdir, recursive = TRUE))

  tmpfile <- file.path(tmpdir, basename(tempfile(fileext = ".html")))
  save_html(x, tmpfile)

  file_out <- paste0(options$label, ".png")

  {
    b$Page$navigate(paste0("file://", tmpfile))
    b$Page$loadEventFired()
  }

  b$Runtime$evaluate(
    "if ($('.dropdown-toggle').length > 0) $('.dropdown-toggle').dropdown('toggle')"
  )
  b$screenshot(filename = here::here("man/figures/", file_out))
}

knitr::opts_chunk$set(eval = TRUE, render = render_screenshot)
```

## A basic example

This first example creates a simple tabbed navigation container with two tabs.
The tab name and the content of each tab are specified in the `nav()` calls
and `navs_tab()` creates the tabbed navigation around these two tabs.

```{r navs_tab-basic, fig.alt = "Screenshot of a basic navs_tab() example."}
library(htmltools)

navs_tab(
  nav(title = "One", p("First tab content.")),
  nav(title = "Two", p("Second tab content."))
)
```

In the rest of the examples, we'll include links among the tabs (or pills) in the navigation controls.

```{r echo = TRUE, eval = TRUE, results="hide"}
link_shiny <- tags$a(shiny::icon("github"), "Shiny", href = "https://github.com/rstudio/shiny", target = "_blank")
link_posit <- tags$a(shiny::icon("r-project"), "Posit", href = "https://posit.co", target = "_blank")
```

## `navs_tab()`

You can fully customize the controls in the navigation component.
In this example, we've added a direct link to the Shiny repository using `nav_item()`.
We've also included a dropdown menu using `nav_menu()`
containing an option to select a third tab panel
and another direct link to Posit's website.
Finally, we've separated the primary tabs on the left
from the direct link and dropdown menu on the right
using `nav_spacer()`.

```{r navs_tab}
navs_tab(
  nav(title = "One", p("First tab content.")),
  nav(title = "Two", p("Second tab content.")),
  nav_spacer(),
  nav_item(link_shiny),
  nav_menu(
    title = "Other links",
    align = "right",
    nav("Three", p("Third tab content")),
    nav_item(link_posit)
  )
)
```

## `navs_pill()`

`navs_pill()` creats a navigation container that behaves exactly like `navs_tab()`,
but the tab toggles are _pills_ or button-shaped.

```{r navs_pill}
navs_pill(
  nav(title = "One", p("First tab content.")),
  nav(title = "Two", p("Second tab content.")),
  nav_spacer(),
  nav_item(link_shiny),
  nav_menu(
    title = "Other links",
    align = "right",
    nav("Three", p("Third tab content")),
    nav_item(link_posit)
  )
)
```

## `navs_tab_card()`

The tabbed navigation container can also be used in a `card()` component
thanks to `navs_tab_card()`.
Learn more about this approach in the
[article about Cards](https://pkgs.rstudio.com/bslib/articles/cards.html).

```{r navs_tab_card}
navs_tab_card(
  nav(title = "One", p("First tab content.")),
  nav(title = "Two", p("Second tab content.")),
  nav_spacer(),
  nav_item(link_shiny),
  nav_menu(
    title = "Other links",
    align = "right",
    nav("Three", p("Third tab content")),
    nav_item(link_posit)
  )
)
```

## `navs_pill_card()`

Similar to `navs_pill()`,
`navs_pill_card()` provides a pill-shaped variant to `navs_tab_card()`.

```{r navs_pill_card}
navs_pill_card(
  nav(title = "One", p("First tab content.")),
  nav(title = "Two", p("Second tab content.")),
  nav_spacer(),
  nav_item(link_shiny),
  nav_menu(
    title = "Other links",
    align = "right",
    nav("Three", p("Third tab content")),
    nav_item(link_posit)
  )
)
```

## `navs_pill_list()`

Furthermore, `navs_pill_list()` creates a vertical list of navigation controls
adjacent to, rather than on top of,
the tab content panels.

```{r navs_pill_list}
navs_pill_list(
  nav(title = "One", p("First tab content.")),
  nav(title = "Two", p("Second tab content.")),
  nav_spacer(),
  nav_item(link_shiny),
  nav_menu(
    title = "Other links",
    align = "right",
    nav("Three", p("Third tab content")),
    nav_item(link_posit)
  )
)
```

## `page_navbar()`

Finally, `page_navbar()` provides full-page navigation container
similar to `navs_tab()` but where each `nav()` is treated as a full page of content
and the navigation controls appear in a top-level navigation bar.

```{r page_navbar}
page_navbar(
  title = "My App",
  bg = "#0062cc",
  nav(title = "One", p("First page content.")),
  nav(title = "Two", p("Second page content.")),
  nav_spacer(),
  nav_item(link_shiny),
  nav_menu(
    title = "Other links",
    align = "right",
    nav("Three", p("Third page content.")),
    nav_item(link_posit)
  )
)
```
