```{r setup, include = FALSE}
library(bslib)

knitr::opts_chunk$set(
  ## Screenshots are created during `devtools::document()` in CI
  ## or set to `TRUE` below to test or update locally
  # .update_screenshot = TRUE,
  "fig.width" = 8,
  "fig.height" = 3,
  "dpi" = 100
)
```

```{r child="0-screenshot-examples.Rmd"}
```

### A basic example

This first example creates a simple tabbed navigation container with two tabs.
The tab name and the content of each tab are specified in the `nav_panel()` calls
and `navset_tab()` creates the tabbed navigation around these two tabs.

```{r navset_tab-basic, fig.alt = "Screenshot of a basic navset_tab() example."}
library(htmltools)

navset_tab(
  nav_panel(title = "One", p("First tab content.")),
  nav_panel(title = "Two", p("Second tab content."))
)
```

In the rest of the examples, we'll include links among the tabs (or pills) in the navigation controls.

```{r echo = TRUE, eval = TRUE, results="hide"}
link_shiny <- tags$a(shiny::icon("github"), "Shiny", href = "https://github.com/rstudio/shiny", target = "_blank")
link_posit <- tags$a(shiny::icon("r-project"), "Posit", href = "https://posit.co", target = "_blank")
```

### `navset_tab()`

You can fully customize the controls in the navigation component.
In this example, we've added a direct link to the Shiny repository using `nav_item()`.
We've also included a dropdown menu using `nav_menu()`
containing an option to select a third tab panel
and another direct link to Posit's website.
Finally, we've separated the primary tabs on the left
from the direct link and dropdown menu on the right
using `nav_spacer()`.

```{r navset_tab}
navset_tab(
  nav_panel(title = "One", p("First tab content.")),
  nav_panel(title = "Two", p("Second tab content.")),
  nav_panel(title = "Three", p("Third tab content")),
  nav_spacer(),
  nav_menu(
    title = "Links",
    nav_item(link_shiny),
    nav_item(link_posit)
  )
)
```

### `navset_pill()`

`navset_pill()` creates a navigation container that behaves exactly like `navset_tab()`,
but the tab toggles are _pills_ or button-shaped.

```{r navset_pill}
navset_pill(
  nav_panel(title = "One", p("First tab content.")),
  nav_panel(title = "Two", p("Second tab content.")),
  nav_panel(title = "Three", p("Third tab content")),
  nav_spacer(),
  nav_menu(
    title = "Links",
    nav_item(link_shiny),
    nav_item(link_posit)
  )
)
```

### `navset_underline()`

`navset_underline()` creates a navigation container that behaves exactly like `navset_tab()` and `navset_pill()`, but the active/focused navigation links are styled with an underline.

```{r navset_underline}
navset_underline(
  nav_panel(title = "One", p("First tab content.")),
  nav_panel(title = "Two", p("Second tab content.")),
  nav_panel(title = "Three", p("Third tab content")),
  nav_spacer(),
  nav_menu(
    title = "Links",
    nav_item(link_shiny),
    nav_item(link_posit)
  )
)
```

### `navset_card_tab()`

The tabbed navigation container can also be used in a `card()` component
thanks to `navset_card_tab()`.
Learn more about this approach in the
[article about Cards](https://pkgs.rstudio.com/bslib/articles/cards.html),
including how to add [a shared sidebar](https://pkgs.rstudio.com/bslib/articles/sidebars.html#multi-page-layout)
to all tabs in the card
using the `sidebar` argument of `navset_card_tab()`.

```{r navset_tab_card}
navset_card_tab(
  nav_panel(title = "One", p("First tab content.")),
  nav_panel(title = "Two", p("Second tab content.")),
  nav_panel(title = "Three", p("Third tab content")),
  nav_spacer(),
  nav_menu(
    title = "Links",
    nav_item(link_shiny),
    nav_item(link_posit)
  )
)
```

### `navset_card_pill()`

Similar to `navset_pill()`,
`navset_card_pill()` provides a pill-shaped variant to `navset_card_tab()`.
You can use the `placement` argument to position the navbar
`"above"` or `"below"` the card body.

```{r navset_card_pill}
navset_card_pill(
  placement = "above",
  nav_panel(title = "One", p("First tab content.")),
  nav_panel(title = "Two", p("Second tab content.")),
  nav_panel(title = "Three", p("Third tab content")),
  nav_spacer(),
  nav_menu(
    title = "Links",
    nav_item(link_shiny),
    nav_item(link_posit)
  )
)
```

### `navset_card_underline()`

`navset_card_underline()` provides a card-based variant of `navset_underline()`.

```{r navset_card_underline}
navset_card_underline(
  nav_panel(title = "One", p("First tab content.")),
  nav_panel(title = "Two", p("Second tab content.")),
  nav_panel(title = "Three", p("Third tab content")),
  nav_spacer(),
  nav_menu(
    title = "Links",
    nav_item(link_shiny),
    nav_item(link_posit)
  )
)
```

### `navset_pill_list()`

Furthermore, `navset_pill_list()` creates a vertical list of navigation controls
adjacent to, rather than on top of, the tab content panels.

```{r navset_pill_list}
navset_pill_list(
  nav_panel(title = "One", p("First tab content.")),
  nav_panel(title = "Two", p("Second tab content.")),
  nav_panel(title = "Three", p("Third tab content")),
  nav_spacer(),
  nav_menu(
    title = "Links",
    nav_item(link_shiny),
    nav_item(link_posit)
  )
)
```

### `page_navbar()`

Finally, `page_navbar()` provides full-page navigation container
similar to `navset_underline()` but where each `nav_panel()` is treated as a full page of content
and the navigation controls appear in a top-level navigation bar.
Note also that the underline styling can be removed via the `underline` argument.

```{r page_navbar}
page_navbar(
  title = "My App",
  bg = "#0062cc",
  underline = TRUE,
  nav_panel(title = "One", p("First tab content.")),
  nav_panel(title = "Two", p("Second tab content.")),
  nav_panel(title = "Three", p("Third tab content")),
  nav_spacer(),
  nav_menu(
    title = "Links",
    align = "right",
    nav_item(link_shiny),
    nav_item(link_posit)
  )
)
```
