---
title: "Sidebars"
output:
  html_document:
    theme:
      version: 5
resource_files:
  - examples
---

```{r setup, include=FALSE}
source(
  rprojroot::find_package_root_file("vignettes/articles/_common.R")
)

describe_layout_function <- function(name, description, element = NULL) {
  name <- knitr::combine_words(sprintf("<code>%s</code>", name),  and = " or ")

  if (is.null(element)) {
    element <- card_body(
      class = "d-flex flex-column align-items-center",
      div(style = "width: 6em; height: 6em; background-color: gray")
    )
  }

  card(
    card_header(class = "bg-info text-center", HTML(name)),
    element,
    card_footer(class = "bg-white text-center", description)
  )
}
```

```{scss, echo = FALSE}
.plotly {
  height: 250px !important;
  .modebar-container {
    display: none;
  }
}

.section {
  margin-top: 2rem;
}

.card-header.bg-info code {
  background: none;
  color: hsl(209deg, 85%, 25%);
}
```

::: lead
Sidebar layouts in web interfaces allow your users to easily access filters, settings and other inputs alongside the interactive features they control.
In the [Getting Started with dashboards](../dashboards) article, we covered "page-level" sidebar layouts via the `page_sidebar()` and `page_navbar()` functions. In this article, we'll explore the full range of sidebar layouts available in bslib.
:::

## Overview

There are three main types of sidebar layouts: floating, filling, and multi-page/tab.

:::: {.row .mt-3}
### Floating layout

Use `layout_sidebar()` to create a sidebar layout that can go anywhere on any page. This layout approach is great for visually grouping together semantically related inputs and output(s). It can also be paired with  a `card()` to leverage `full_screen` expansion, add a header/footer, and more.

::: {.col-sm-12 .col-lg-6}

<details>
<summary>Show code</summary>
```r
layout_sidebar(
  sidebar = sidebar("Sidebar"),
  "Main contents"
)
```
</details>

```{r preview-layout-sidebar, echo = FALSE}
describe_layout_function(
  "layout_sidebar()",
  "A sidebar layout.",
  card_body(
    class = "p-2",
    height = 200,
    layout_sidebar(
      sidebar = sidebar("Sidebar", width = "33%", class = "bg-light"),
      "Main contents"
    ) |>
      tagAppendAttributes(class = "border rounded-top")
  )
)
```
:::

::: {.col-sm-12 .col-lg-6}

<details>
<summary>Show code</summary>
```r
card(
  full_screen = TRUE,
  card_header("Title"),
  layout_sidebar(
    sidebar = sidebar("Sidebar"),
    "Main contents"
  )
)
```
</details>

```{r preview-layout-sidebar-card, echo = FALSE}
describe_layout_function(
  "layout_sidebar() in card()",
  "A sidebar layout within a card.",
  card_body(
    class = "p-2",
    height = 200,
    card(
      full_screen = TRUE,
      card_header("Title"),
      layout_sidebar(
        sidebar = sidebar("Sidebar", width = "33%", class = "bg-light"),
        "Main contents"
      )
    )
  )
)
```
:::
::::


:::: {.row .mt-5}
### Filling layout

::: {.col-sm-12 .col-lg-6}
In the [Getting Started with dashboards](../dashboards) article, we saw how `page_sidebar()` yields a sidebar layout that fills the page. Underneath the hood, `page_sidebar()` is just a simple wrapper around `page_fillable()` and `layout_sidebar()`. Understanding this unlocks the potential to have (any number of) sidebar layouts within a filling layout.
:::

::: {.col-sm-12 .col-lg-6}

<details>
<summary>Show code</summary>
```r
page_fillable(
  layout_sidebar(
    sidebar = sidebar("Sidebar area"),
    "Main area"
  )
)
```
</details>

```{r preview-page-fillable, echo = FALSE}
describe_layout_function(
  "layout_sidebar() in page_fillable()",
  "A sidebar that fills the page.",
  card_body(
    class = "p-0",
    height = 200,
    page_fillable(
      layout_sidebar(
        sidebar = sidebar("Sidebar", width = "33%", class = "bg-light"),
        "Main contents"
      )
    )
  ) |>
  tagAppendAttributes(class = "flex-row", .cssSelector = ".navbar-nav") |>
  tagAppendAttributes(class = "d-none", .cssSelector = ".navbar-header") |>
  tagAppendAttributes(class = "navbar-dark", .cssSelector = "nav")
)
```
:::
::::


:::: {.row .mt-5}

### Multi-page layout

For a multi-page (or multi-tab) layout, use the `sidebar` argument of `page_navbar()` (or `navset_card_tab()`). In this case, we get a sidebar that not only fills the page, but that _same sidebar remains visible on every page/tab_. Later on, we'll explore how to put [multiple, varied, layouts on different pages](#multi-page-example); but also keep in mind, if it is actually desirable to have the _same sidebar_ on every page, it often helps to [hide/show sidebar contents on certain pages](#conditional-contents) via `conditionalPanel()`.

::: {.col-sm-12 .col-lg-6}

<details>
<summary>Show code</summary>
```r
page_navbar(
  sidebar = sidebar("Sidebar"),
  nav_panel("Page 1", "Page 1 content"),
  nav_panel("Page 2", "Page 2 content")
)
```
</details>

```{r preview-page-navbar, echo = FALSE}
describe_layout_function(
  "page_navbar()",
  "A sidebar shared across multiple pages.",
  card_body(
    class = "p-0",
    style = "gap:0;",
    height = 200,
    id = "page_navbar_example_card",
    navset_bar(
      collapsible = FALSE,
      sidebar = sidebar("Shared sidebar", width = "33%", class = "bg-light"),
      nav_panel("Page 1", "Page 1 content"),
      nav_panel("Page 2", "Page 2 content")
    ),
    tags$script(HTML("$(document).ready(function() {
      // pkgdown activates headroom on the navset_bar() container and ends up
      // auto-hiding our navbar, so this destroys the unwanted headroom instance
      var demo_nav = $('#page_navbar_example_card nav.navbar');
      setTimeout(function() {
        if (demo_nav.headroom) demo_nav.headroom('destroy')
      }, 100);
    });"))
  ) |>
  tagAppendAttributes(class = "flex-row", .cssSelector = ".navbar-nav") |>
  tagAppendAttributes(class = "d-none", .cssSelector = ".navbar-header") |>
  tagAppendAttributes(class = "navbar-dark", .cssSelector = "nav")
)
```
:::

::: {.col-sm-12 .col-lg-6}

<details>
<summary>Show code</summary>
```r
navset_card_tab(
  sidebar = sidebar("Sidebar"),
  nav_panel("Tab 1", "Tab 1 content"),
  nav_panel("Tab 2", "Tab 2 content")
)
```
</details>

```{r preview-navs-tab-card, echo = FALSE}
describe_layout_function(
  "navset_card_tab()",
  "A sidebar shared across multiple tabs.",
  card_body(
    class = "p-2",
    height = 200,
    navset_card_tab(
      title = "Tab Card",
      sidebar = sidebar("Shared sidebar", width = "33%", class = "bg-light"),
      nav_panel("Tab 1", "Tab 1 content"),
      nav_panel("Tab 2", "Tab 2 content")
    )
  )
)
```
:::
::::


## A real example

Now that we've enumerated bslib's sidebar layout options, lets use some real data[^real-data] to create some real inputs and outputs, and explore some additional features of sidebar layouts.

In a Shiny app[^shiny-id], you'll probably want to use inputs like `selectInput()`, `sliderInput()`, etc., in the sidebar, but because you're reading this article in a static website, we'll use [crosstalk](https://rstudio.github.io/crosstalk/) input widgets.

[^real-data]: Our "real data" is just a 1,000 rows randomly sampled from `{ggplot2}`'s `diamonds` data as well as `{leaflet}`'s `quakes`.
[^shiny-id]: In a Shiny app, we also recommend you add an `id` to the `sidebar()` so that you can reactively read/update whether the sidebar is open/closed.

### Setup code

Throughout this section, we'll make repeated use of the following widgets from `{plotly}` and `{leaflet}`. The details on how these widgets work alongside `{crosstalk}` to create linked views isn't important for understanding sidebar layouts, but do keep in mind this will give us a list of `filters` and `plots` (views of the `diamonds` dataset), as well as `map_filter` and `map_quakes` (views of the `quakes` dataset).

<details>
<summary>Show code</summary>

```{r}
library(bslib)
library(shiny)
library(crosstalk)
library(plotly)
library(leaflet)

# Creates the "filter link" between the controls and plots
dat <- SharedData$new(dplyr::sample_n(diamonds, 1000))

# Sidebar elements (e.g., filter controls)
filters <- list(
  filter_select("cut", "Cut", dat, ~cut),
  filter_select("color", "Color", dat, ~color),
  filter_select("clarity", "Clarity", dat, ~clarity)
)

# plotly visuals
plots <- list(
  plot_ly(dat) |> add_histogram(x = ~price),
  plot_ly(dat) |> add_histogram(x = ~carat),
  plot_ly(dat) |> add_histogram(x = ~cut, color = ~clarity)
)
plots <- lapply(plots, \(x) config(x, displayModeBar = FALSE))

# map filter and visual
quake_dat <- SharedData$new(quakes)
map_filter <- filter_slider("mag", "Magnitude", quake_dat, ~mag)
map_quakes <- leaflet(quake_dat) |>
  addTiles() |>
  addCircleMarkers()
```

</details>

### Hello `layout_sidebar()` {#hello-sidebar}

`layout_sidebar()` behaves a lot like a [card](../cards). For example, when used inside `page_fillable()` they'll also grow/shrink to fit the page (because they default to `fill = TRUE`). They also default to `fillable = TRUE` which allows fill items in the main content area (e.g., `plots[[1]]`) to also grow/shrink to fit their container. They also behave a lot like a `card_body()` in that they can be put directly inside a `card()` (which is useful for adding a header/footer, `full_screen = TRUE`, etc.).

```{r hello-layout-sidebar, as_iframe = TRUE, resizable = TRUE}
sidebar_diamonds <- layout_sidebar(
  sidebar = filters[[1]],
  plots[[1]]
)

sidebar_quakes <- layout_sidebar(
  sidebar = map_filter,
  map_quakes
)

page_fillable(
  sidebar_diamonds,
  card(
    card_header("Earthquakes"),
    sidebar_quakes
  )
)
```

:::: {.row .my-3}
::: {.col-sm-12 .col-lg-6}
::: {.callout .callout-note}
<h3 data-toc-skip>Resizable example</h3>

The example above is resizable. Try using the handle in the lower-right corner to change the "window" size and notice how the plot grow/shrink to fit the window (because of `fillable = TRUE`).
:::
:::

::: {.col-sm-12 .col-lg-6 .mt-auto .mb-auto}
::: {.callout .callout-note}
<h3 data-toc-skip>Filling layouts</h3>

To learn more about how fillable containers and fill items work, see the article on [filling layouts](../filling).
:::
:::
::::

### Multi-page varied layout {#multi-page-example}

As we covered in [Getting Started with dashboards](../dashboards), the `sidebar` argument of `page_navbar()` puts a sidebar _on each page_ that fills the window. However, sometimes it's better that only particular pages have such a sidebar layout. To acheive this, just provide a `layout_sidebar()` as a "root" element of a `fillable` page.

For example, let's put a "page-level" sidebar on a page dedicated to Earthquakes, and then put multiple sidebar layouts on a page dedicated to Diamonds (one for each plot). In this case, we've only allowed the Earthquakes page to be `fillable` since there are multiple plots on the Diamonds page (you could also keep the Diamonds page `fillable` an put a `min_height` on the cards to prevent them from shrinking too much).

```{r multi-page, as_iframe = TRUE, resizable = TRUE}
page_navbar(
  title = "Sidebar demo",
  fillable = "Earthquakes",
  nav_panel("Earthquakes", sidebar_quakes),
  nav_panel(
    "Diamonds",
    Map(
      function(filter, plot) {
        card(
          full_screen = TRUE,
          layout_sidebar(sidebar = filter, plot)
        )
      },
      filters, plots
    )
  )
)
```


::: {.callout .callout-note}
<h3 data-toc-skip>Multiple tabs</h3>

Just like `page_navbar()`, `navset_card_tab()` also has a `sidebar` argument that puts _the same sidebar_ on each tab. The same approach (i.e., putting a `layout_sidebar()` within each `nav_panel()`) can be used to put different sidebars on different tabs.
:::


### Restricting growth

Just like with [cards](../cards#restricting-growth), when a [filling layout](#hello-sidebar) isn't enforcing the size of the `layout_sidebar()`, it will allow it's contents to decide how big it should be. Thus, if there a large amount of sidebar/main contents, consider specifying a `height` or `max_height` via `card()` (as well as `full_screen = TRUE` to reduce the need for scrolling).

```{r restricting-growth, as_iframe = TRUE, resizable = TRUE}
page_fixed(
  h1("Sidebar demo", class = "lead mt-3"),
  card(
    height = 400,
    full_screen = TRUE,
    layout_sidebar(sidebar = filters, plots)
  ),
  card(
    full_screen = TRUE,
    layout_sidebar(sidebar = map_filter, map_quakes)
  )
)
```


## Shiny

Although sidebars work just fine outside Shiny, using them in Shiny provides a few additional useful features.

### Conditional contents

Sometimes in a multiple page/tab setting, it's useful to have a sidebar on every page/tab, but changes it's contents based on which page/tab is active.[^4] Thanks to `conditionalPanel()`, this can be done fairly easily in a Shiny app with `page_navbar()` (or in `navset_card_tab()`/`navset_tab_pill()`). The trick is to provide an `id` to the `page_navbar()` and then reference that `id` in the `conditionalPanel()`:

[^4]: If the controls depend on some other application state, you'll need to use `uiOutput()` to fill the contents of a `sidebar()`).

:::: row
::: {.col-sm-12 .col-lg-6}
```{r, eval = FALSE}
shinyApp(
  page_navbar(
    title = "Conditional sidebar",
    id = "nav",
    sidebar = sidebar(
      conditionalPanel(
        "input.nav === 'Page 1'",
        "Page 1 sidebar"
      ),
      conditionalPanel(
        "input.nav === 'Page 2'",
        "Page 2 sidebar"
      )
    ),
    nav_panel("Page 1", "Page 1 contents"),
    nav_panel("Page 2", "Page 2 contents")
  ),
  server = function(...) {
    # no server logic required
  }
)
```
:::

::: {.col-sm-12 .col-lg-6 .mt-auto .mb-auto}
```{r shiny-dynamic-sidebar, echo = FALSE, as_iframe = TRUE}
page_navbar(
  title = "Conditional sidebar",
  id = "nav",
  collapsible = FALSE,
  sidebar = sidebar(
    open = TRUE,
    conditionalPanel(
      "input.nav === 'Page 1'",
      "Page 1 sidebar"
    ),
    conditionalPanel(
      "input.nav === 'Page 2'",
      "Page 2 sidebar"
    )
  ),
  nav_panel("Page 1", "Page 1 contents"),
  nav_panel("Page 2", "Page 2 contents"),
  footer = htmltools::tags$script(htmltools::HTML("
var $page1,$page2;
$(document).ready(function() {
  $page1 = $(`[data-display-if*='Page 1']`);
  $page2 = $(`[data-display-if*='Page 2']`);
  $page2.hide();
});

$('#nav').on('shown.bs.tab', function(ev) {
  const page = ev.target.dataset.value;
  page === 'Page 2' ? $page1.hide() : $page1.show();
  page === 'Page 1' ? $page2.hide() : $page2.show();
});
"
  ))
)
```
:::
::::


### Reactive updates

To programmatically update (and/or re-actively read) the open/closed state of a `sidebar()`, provide an `id` and reference that `id` in your server code. Here we reference use the `id` to programmatically open the sidebar on the 2nd page.


:::: row
::: {.col-sm-12 .col-lg-6}
```{r eval = FALSE}
library(shiny)

ui <- page_navbar(
  title = "Sidebar updates",
  id = "nav",
  sidebar = sidebar(
    id = "sidebar",
    open = FALSE,
    "Sidebar"
  ),
  nav_panel("Page 1", "Sidebar closed. Go to Page 2 to open."),
  nav_panel("Page 2", "Sidebar open. Go to Page 1 to close.")
)

server <- function(input, output) {
  observe({
    sidebar_toggle(
      id = "sidebar",
      open = input$nav == "Page 2"
    )
  })
}

shinyApp(ui, server)
```
:::

::: {.col-sm-12 .col-lg-6 .mt-auto .mb-auto}
```{r shiny-sidebar-state, echo = FALSE, as_iframe = TRUE}
page_navbar(
  title = "Sidebar updates",
  id = "nav",
  collapsible = FALSE,
  sidebar = sidebar(id = "sidebar", open = FALSE, "Sidebar"),
  nav_panel("Page 1", "Sidebar closed. Go to Page 2 to open."),
  nav_panel("Page 2", "Sidebar open. Go to Page 1 to close."),
  footer = htmltools::tags$script(htmltools::HTML("
$('#nav').on('shown.bs.tab', function(ev) {
  const page = ev.target.dataset.value;
  const layout = document.getElementById('sidebar').parentElement;
  const open = page === 'Page 2' ? 'open' : 'close';
  bslib.Sidebar.getInstance(layout).toggle(open);
});"
  ))
)
```
:::
::::




## Accordions

All sidebars have special treatment for accordions.
When an `accordion()` appears directly within a `sidebar()` (as an immediate child of the sidebar), the accordion panels will render flush to the sidebar, providing a convenient way to group multiple related input controls under a collapsible section.

::: {.callout .callout-note}
<h3 data-toc-skip>Setup code</h3>

This example depends on objects from the [setup code section](#setup-code).
:::

```{r}
accordion_filters <- accordion(
  accordion_panel(
    "Dropdowns", icon = bsicons::bs_icon("menu-app"),
    !!!filters
  ),
  accordion_panel(
    "Numerical", icon = bsicons::bs_icon("sliders"),
    filter_slider("depth", "Depth", dat, ~depth),
    filter_slider("table", "Table", dat, ~table)
  )
)

card(
  card_header("Groups of diamond filters"),
  layout_sidebar(
    sidebar = accordion_filters,
    plots[[1]]
  )
)
```


## Nested sidebars

It's possible to nest sidebar layouts, which means you can effectively have any number of left and/or right sidebars in a given layout. When doing this, you'll want the main content area of every `layout_sidebar()` that contains a `layout_sidebar()` to be fillable and have zero padding (`class = "p-0"`).

```{r page-fill-double, as_iframe = TRUE}
page_fillable(
  h1("Left and right sidebar", class = "px-3 my-3"),
  layout_sidebar(
    sidebar = sidebar("Left sidebar"),
    layout_sidebar(
        sidebar = sidebar("Right sidebar", position = "right", open = FALSE),
        "Main contents",
        border = FALSE
    ),
    border_radius = FALSE,
    fillable = TRUE,
    class = "p-0"
  )
)
```

## Styling and customization

In the above sections we've focused primarily on the variety of interface layouts where sidebars can be used.
Along the way, we've touched on a few of the named arguments of `sidebar()` and `layout_sidebar()` that are helpful for customizing the styling and behavior of both the sidebar and main content areas.
However, there are a handful of other arguments to further customize the look and feel if the sidebar layout.

Both `sidebar()` and `layout_sidebar()` allow for a specific background color (via `bg`), which is applied to the sidebar area and main content area respectively.
When `bg` is provided, bslib automatically provides a high-contrast foreground color to ensure readability (but a `fg` color may also be provided).
Both functions also include a `class` argument that works well with [Bootstrap utility `class`es](https://getbootstrap.com/docs/5.3/utilities/spacing/) and a `style` argument for inline styles.

Be aware that in `layout_sidebar()`, `bg`, `class` and `style` attributes are applied to the main content area's container and _not_ the overall layout container.
To add additional classes to the layout container, use `htmltools::tagAppendAttributes()`.
Also note that `layout_sidebar()` derives some of it's default style from Bootstrap CSS variables (e.g., `--bs-card-border-color`), which enables theming at the component-level ([theming via `bs_theme()`](../bs5-variables) works on the page-level).

The following example combines all of these concepts to create sidebar with a dark background.
Utility classes are used to make the sidebar text monospace and bold, and we used `tagAppendAttributes()` to tweak the border color of the sidebar layout to match the sidebar background.

```{r}
library(htmltools)
library(leaflet)

squake <- SharedData$new(quakes)

container <- layout_sidebar(
  class = "p-0",
  sidebar = sidebar(
    title = "Earthquakes off Fiji",
    bg = "#1E1E1E",
    width = "35%",
    class = "fw-bold font-monospace",
    filter_slider("mag", "Magnitude", squake, ~mag)
  ),
  leaflet(squake) |> addTiles() |> addCircleMarkers()
)

tagAppendAttributes(container, style = css("--bs-card-border-color" = "#1E1E1E"))
```
